---
title: "Amaizing  Baseline Survey_Final-rp_risk"
author: "Njambi"
format: html
editor: visual
---

## 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Clear the environment

```{r}

rm(list = ls())

```

## Loading necessary libraries

```{r, include=FALSE}

library(tidyverse) # includes the dplyr, ggplot2, tidyr, readr, stringr, forcats, tibble packages
library(readxl)
library(Hmisc)
library(labelled)
library(magrittr)
library(datawizard)
library(gtsummary)
library(gt)
library(kableExtra)
library(flextable)
library(splitstackshape)
library(lubridate)
library(foreign)
library(haven)


```

## Set the working directory

```{r}
# Set working directory

setwd("C:\\Users\\awanjau\\OneDrive - CGIAR\\Alliance\\Alliance_2024\\AMAZING\\AmazingBaselineDataCleaning\\Amaizing  Baseline Survey_Final-rp_risk")
      
```

## Loading data

```{r}
# Read a csv file

aMaizingBL_risks_path <- "C:\\Users\\awanjau\\OneDrive - CGIAR\\Alliance\\Alliance_2024\\AMAZING\\aMaizingBaseline_Files\\Amaizing  Baseline Survey_Final-rp_risk.csv"

aMaizingBL_risks <- read_csv(aMaizingBL_risks_path)

head(aMaizingBL_risks)
```

```{r}
# Check number of rows and columns in data

dim(aMaizingBL_risks)

# Print column names

names(aMaizingBL_risks)
```

```{r}
class(aMaizingBL_risks$riskindex1)

sort(unique(aMaizingBL_risks$riskindex1), na.last = TRUE)
```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    riskindex1 = as.integer(riskindex1)
  )

class(aMaizingBL_risks$riskindex1)
```

```{r}
class(aMaizingBL_risks$riskindex2)

sort(unique(aMaizingBL_risks$riskindex2), na.last = TRUE)
```

```{r}

risk_levels <- c(
  
 "Drought",	"Excessive Rainfall",	"Pests",	"Diseases",	"Floods",	"Animals",	"Fire",	"Bird",	"Frost",	"Hail", 	"Theft", 	"High input price",	"Input shortages",	"Commodity price fluctuations",	"Lack of market",	"Other, Specify"

                         )  

aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    riskindex2 = factor(
      riskindex2, 
      levels = c(1:15, -222), 
      labels = risk_levels
    )
  
  ) 

class(aMaizingBL_risks$riskindex2)
unique(aMaizingBL_risks$riskindex2)
```

```{r}
class(aMaizingBL_risks$riskname)

#sort(unique(aMaizingBL_risks$riskname), na.last = TRUE)
```

```{r}

aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    riskname = factor(riskname)
  
  ) 

class(aMaizingBL_risks$riskname)
#unique(aMaizingBL_risks$riskname)

# Not cleaned; includes text from "Other, Specify"
```

```{r}
class(aMaizingBL_risks$f31)

sort(unique(aMaizingBL_risks$f31), na.last = TRUE)
```

1 Every year 2 Every two years 3 Every three years 4 More than three years

```{r}

risk_freq_levels <- c(
  
 "Every year",	"Every two years",	"Every three years",	"More than three years"

                         )  

aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f31 = factor(
      f31, 
      levels = c(1:4), 
      labels = risk_freq_levels ,
      ordered = TRUE
    )
  
  ) 

class(aMaizingBL_risks$f31)
unique(aMaizingBL_risks$f31)
```

***Reverse code f31 and convert to ordinal***

```{r}

aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f31_reversed = factor(
      f31, 
     # levels = rev(levels(f31)),  # Reverse the levels while keeping the labels intact
      levels = rev(levels(f31)) ,  # Reverse the levels
     # labels = rev(risk_freq_levels),  # Reverse the labels to match reversed levels
      ordered = TRUE   
    )
  )

class(aMaizingBL_risks$f31_reversed)
unique(aMaizingBL_risks$f31_reversed)
```

```{r}
class(aMaizingBL_risks$f32)

sort(unique(aMaizingBL_risks$f32), na.last = TRUE)
```

1 Slightly (1 -19%) 2 Moderately ( 20%-49%) 3 Severe (50%-74%) 4 Extremely (75%-100%)

```{r}

severity_levels <- c(
  
 "Slightly (1 -19%)",	"Moderately ( 20%-49%)",	"Severe (50%-74%)",	"Extremely (75%-100%)"

                         )  

aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f32 = factor(
      f32, 
      levels = c(1:4), 
      labels = severity_levels ,
      ordered = TRUE
    )
  
  ) 

class(aMaizingBL_risks$f32)
unique(aMaizingBL_risks$f32)
```

```{r}
class(aMaizingBL_risks$f33)

#sort(unique(aMaizingBL_risks$f33), na.last = TRUE)

# Character
# Multi-select
```

```{r}
class(aMaizingBL_risks$f33_1)

sort(unique(aMaizingBL_risks$f33_1), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33_1 = as.integer(f33_1)
  )

class(aMaizingBL_risks$f33_1)
```

```{r}
class(aMaizingBL_risks$f33_2)

sort(unique(aMaizingBL_risks$f33_2), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33_2 = as.integer(f33_2)
  )

class(aMaizingBL_risks$f33_2)
```

```{r}
class(aMaizingBL_risks$f33_3)

sort(unique(aMaizingBL_risks$f33_3), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33_3 = as.integer(f33_3)
  )

class(aMaizingBL_risks$f33_3)
```

```{r}
class(aMaizingBL_risks$f33_4)

sort(unique(aMaizingBL_risks$f33_4), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33_4 = as.integer(f33_4)
  )

class(aMaizingBL_risks$f33_4)
```

```{r}
class(aMaizingBL_risks$f33_5)

sort(unique(aMaizingBL_risks$f33_5), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33_5 = as.integer(f33_5)
  )

class(aMaizingBL_risks$f33_5)
```

```{r}
class(aMaizingBL_risks$f33_6)

sort(unique(aMaizingBL_risks$f33_6), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33_6 = as.integer(f33_6)
  )

class(aMaizingBL_risks$f33_6)
```

```{r}
class(aMaizingBL_risks$f33_7)

sort(unique(aMaizingBL_risks$f33_7), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33_7 = as.integer(f33_7)
  )

class(aMaizingBL_risks$f33_7)
```

```{r}
class(aMaizingBL_risks$f33_8)

sort(unique(aMaizingBL_risks$f33_8), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33_8 = as.integer(f33_8)
  )

class(aMaizingBL_risks$f33_8)
```

```{r}
class(aMaizingBL_risks$f33_9)

sort(unique(aMaizingBL_risks$f33_9), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33_9 = as.integer(f33_9)
  )

class(aMaizingBL_risks$f33_9)
```

```{r}
class(aMaizingBL_risks$f33_10)

sort(unique(aMaizingBL_risks$f33_10), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33_10 = as.integer(f33_10)
  )

class(aMaizingBL_risks$f33_10)
```

```{r}
class(aMaizingBL_risks$f33_11)

sort(unique(aMaizingBL_risks$f33_11), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33_11 = as.integer(f33_11)
  )

class(aMaizingBL_risks$f33_11)
```

```{r}
class(aMaizingBL_risks$f33_12)

sort(unique(aMaizingBL_risks$f33_12), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33_12 = as.integer(f33_12)
  )

class(aMaizingBL_risks$f33_12)
```

```{r}
class(aMaizingBL_risks$f33_13)

sort(unique(aMaizingBL_risks$f33_13), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33_13 = as.integer(f33_13)
  )

class(aMaizingBL_risks$f33_13)
```

```{r}
class(aMaizingBL_risks$f33_14)

sort(unique(aMaizingBL_risks$f33_14), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33_14 = as.integer(f33_14)
  )

class(aMaizingBL_risks$f33_14)
```

```{r}
class(aMaizingBL_risks$f33_15)

sort(unique(aMaizingBL_risks$f33_15), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33_15 = as.integer(f33_15)
  )

class(aMaizingBL_risks$f33_15)
```

```{r}
class(aMaizingBL_risks$f33_16)

sort(unique(aMaizingBL_risks$f33_16), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33_16 = as.integer(f33_16)
  )

class(aMaizingBL_risks$f33_16)
```

```{r}
class(aMaizingBL_risks$f33_17)

sort(unique(aMaizingBL_risks$f33_17), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33_17 = as.integer(f33_17)
  )

class(aMaizingBL_risks$f33_17)
```

```{r}
class(aMaizingBL_risks$f33_18)

sort(unique(aMaizingBL_risks$f33_18), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33_18 = as.integer(f33_18)
  )

class(aMaizingBL_risks$f33_18)
```

```{r}
class(aMaizingBL_risks$f33__222)

sort(unique(aMaizingBL_risks$f33__222), na.last = TRUE)

```

```{r}
aMaizingBL_risks <- aMaizingBL_risks %>%
  mutate(
    f33__222 = as.integer(f33__222)
  )

class(aMaizingBL_risks$f33__222)
```

```{r}
class(aMaizingBL_risks$f33_sp)

#sort(unique(aMaizingBL_risks$f33_sp), na.last = TRUE)

# Text data; not cleaned

```

```{r}

column_labels <- c(

  riskindex1 = "Climatic/production risk repeat iteration/index",
  riskindex2 = "Climatic/production risk level/category selected",
  riskname = "Climatic/production risk corresponding to riskindex2 value",
  f31 = "Frequency of occurrence of riskname",
  f31_reversed = "Reverse-coded frequency of occurrence of riskname",
  f32 = "How severe was the damage to the cname (crop name) harvested for the previous season due to riskname?",
  f33 = "Coping strategy used to deal with the loss caused by riskname?",
  f33_1 = "Nothing",
  f33_2 = "Borrow money",
  f33_3 = "Use cash/savings",
  f33_4 = "Use Insurance",
  f33_5 = "Sell stocks of food/crops",
  f33_6 = "Sell animals/farmland/ assets",
  f33_7 = "Invest less in agriculture",
  f33_8 = "Invest less in other businesses/activities",
  f33_9 = "Reduce food expenditure",
  f33_10 = "Reduce non-food expenditures",
  f33_11 = "Postpone expenditures",
  f33_12 = "Work more hours / seek employment",
  f33_13 = "Send children to relatives or friends",
  f33_14 = "Rely on gifts/assistance from friends or relatives",
  f33_15 = "Seek religious/spiritual help / pray",
  f33_16 = "Get help from an NGO",
  f33_17 = "Through savings/credit group",
  f33_18 = "Increased Cost of production",
  f33__222 = "Other (specify)",
  f33_sp = "Please specify the Coping Strategy",
  PARENT_KEY = "Parent key",
  KEY = "Key",
  `SET-OF-rp_risk` = "Climate/production risk repeat group ID"

)

```

```{r}

# Rename columns with characters not supported by stata

aMaizingBL_risks <- aMaizingBL_risks %>%
  
  rename(SET_OF_rp_risk = `SET-OF-rp_risk`
    
         )

```

```{r}
#names(aMaizingBL_risks)

aMaizingBL_risks <- aMaizingBL_risks %>%
  select(1:4, 30, 5:29) %>%
  set_variable_labels(.labels = column_labels)

names(aMaizingBL_risks)
```

Using the function View() to open the new data in a spreadsheet reveals the column names and labels.

```{r}

#View(aMaizingBL_plot_level)

```

## Generate a data dictionary

```{r}
# Data dictionary

aMaizingBL_risks_dict <- look_for(aMaizingBL_risks, labels = TRUE, values = TRUE) %>%
  dplyr::select(-c(missing, value_labels))

  gt(aMaizingBL_risks_dict)

```

## Export data dictionary to CSV

```{r}

readr::write_csv(aMaizingBL_risks_dict, "aMaizingBL_risks_dict.csv")

```

## Export data to CSV

```{r}

readr::write_csv(aMaizingBL_risks, "aMaizingBL_risks.csv")

```

## Save as R data

```{r}

save(aMaizingBL_risks, file = "aMaizingBL_risks.RData")
```

## Export data to SPSS

```{r}

write.foreign(aMaizingBL_risks, "aMaizingBL_risks.sav", "aMaizingBL_risks", package = "SPSS")

```

## Export data to STATA

```{r}

write_dta(aMaizingBL_risks, "aMaizingBL_risks.dta")

```
